Lucid: Lifting-based Uncertain Control Invariant Dynamics

### 2025-07-26 docs/paper/appendix.tex
feat: add theory

\begin{appendices}
    \section{Proofs}
    We have collected a series of proofs of the claims we made in the main text.

    \section{Benchmarks}
    \label{app:benchmarks}
    Complete list of results for the benchmarks presented in \autoref{sec:benchmarks}.
    The legend is as follows:
    \textit{Output} indicates the result produced by the solver (i.e., optimal, unbounded, infeasible or unspecified),
    \textit{Prec} indicates the number of bits used in the last floating-point number representation,
    \textit{Ref} indicates the number of refinements,
    and \textit{time} is the time in seconds.

\end{appendices}

>>>>

% !TEX root =  main.tex
\appendix

\section{Additional Theoretical Details}
\paragraph{RKHS basics.}
A symmetric function $k_\X:\X\times\X\rightarrow\R$ is called a (positive definite) \emph{kernel} (note the distinction from \emph{probability kernels}) if for all $N\in\N_{>0}$ we have $\sum_{i=1}^{N}\sum_{j=1}^{N}a_i a_j\allowbreak k_\X(x_i,x_j) \geq 0$ for $x_1,\ldots,x_N\in\X\subset{\R^n}$ and $a_1,\ldots,a_N\in\R$.
A prominent example is the \emph{squared exponential} (SQExp) kernel \shortcite{Rasmussen2005GP,Kanagawa2018GPvsKernel}:
\begin{equation}
    k_\X(x,x') := \sigma_f^2 \exp\left( -\frac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\quad \Sigma:=\mathrm{diag}(\sigma_l)^{-2},\label{eq:sqexp_kernel}
\end{equation}
with amplitude $\sigma_f^2\geq0$ and lengthscale coefficients $\sigma_l\in\R^n$.
\new{For this work,} we assume that all kernels are bounded on their domain, i.e., $\E_{}[k_\X(x,x)]<\infty$, $x\in\X$.
Given a kernel $k_\X$ on a non-empty set $\X$, there exists a unique corresponding
\emph{reproducing kernel Hilbert space} (RKHS) $\Hilbert_{k_\X}$
of functions $f:\X\rightarrow\R$ equipped with an inner product $\innerH{\cdotx}{\cdotx}{\Hilbert_{k_\X}}$
with the celebrated \emph{reproducing property} such that for any function $f\in\Hilbert_{k_\X}$ and $x\in\X$ we have $f(x)=\innerH{f}{k_\X(\cdotx,x)}{\Hilbert_{k_\X}}$.
Note that $k_\X(\cdotx,x):\X\rightarrow\Hilbert_{k_\X}$ is a real-valued function,
which is also called an implicit \emph{canonical} \emph{embedding} or \emph{feature map} $\phi_\X$
such that $k_\X(x,x')=\innerH{\phi_\X(x)}{\phi_\X(x')}{\Hilbert_{k_\X}}$ for all $x,x'\in\X$.
For an RKHS $\Hilbert_{k_\X}$, we use the associated feature map $\phi_\X$ and kernel $k_\X$ interchangeably for ease of notation and comprehensibility.
The inner product induces the norm $\norm{f}_{\Hilbert_{k_\X}}\!\!\!\!:=\!\!\sqrt{\smash[b]{\innerH{f}{f}{\Hilbert_{k_\X}}}}$ of the RKHS. 
Throughout this paper, we assume that all RKHSs are \emph{separable}.
Refer to the monograph by \citet{Berlinet2004RKHSProbStat} for a comprehensive study on RKHSs.
Given $N$ i.i.d. samples $\hat X_N:=[\hat{x}_i]_{i=1}^N$ with $\hat{x}_i\in\X$, the
\emph{Gram matrix} of $k_\X$ is given by
$\new{K_{\hat{X}}^N}:=[k_\X(\hat{x}_i,\hat{x}_j)]_{i,j=1}^N.$
Furthermore, we define the vector-valued function 
$\new{k_{\hat{X}}^N}(x)  := [k_\X(x,\hat{x}_i)]_{i=1}^N.$

\begin{definition}[Conditional Mean Embedding (CME)]\label{def:condMeanEmbed}
    Given two RKHSs $\Hilbert_{k_\X}$ and $\Hilbert_{k_\Y}$ with the associated kernels $k_\X\colon\X\times\X\to\R$ and $k_\Y\colon\X\times\X\to\R$, the \emph{CME} of a probability kernel $\Tr\colon\X\times\borel{\Y}\rightarrow[0,1]$ is an $X$-measurable random variable taking values in $\Hilbert_{k_\Y}$, given by
    \begin{equation*}
        \cme_{k_\Y|k_\X}(\Tr)(\cdotx) := \E_{\Tr}[\phi_\Y(Y)\mid X=\cdotx].
    \end{equation*}
\end{definition}

    % \section{Proofs}
    % We have collected a series of proofs of the claims we made in the main text.

    % \section{Benchmarks}
    % \label{app:benchmarks}
    % Complete list of results for the benchmarks presented in \autoref{sec:benchmarks}.
    % The legend is as follows:
    % \textit{Output} indicates the result produced by the solver (i.e., optimal, unbounded, infeasible or unspecified),
    % \textit{Prec} indicates the number of bits used in the last floating-point number representation,
    % \textit{Ref} indicates the number of refinements,
    % and \textit{time} is the time in seconds.

### 2025-07-26 docs/paper/figures/architecture.tex
feat: add theory

\begin{figure}
    \centering
    \begin{tikzpicture}[node distance=0.6cm and 0.6cm,
            input-block/.style={rectangle, draw, fill=red!10, text width=2.5cm, minimum height=1.2cm, align=center, rounded corners},
            output-block/.style={rectangle, draw, fill=green!10, text width=2.5cm, minimum height=1.2cm, align=center, rounded corners},
            block/.style={rectangle, draw, fill=blue!10, text width=2.0cm, minimum height=1.2cm, align=center, rounded corners},
            arrow/.style={thick,->,>=Stealth},
            tool/.style={rectangle, draw, fill=magenta!10, text width=3cm, minimum height=0.8cm, align=center, rounded corners},
            inv/.style={}
        ]

        % Input block
        \node[input-block] (input) {Configuration};

        % Lucid blocks
        \node[block, below=of input] (estimator) {Estimator};
        \node[block, right=of estimator] (tuner) {Tuner};
        \node[block, left=of estimator] (sets) {Sets};
        \node[block, below=of estimator] (feature) {Feature map / Barrier template};
        \node[block, left=of feature] (optimizer) {Optimizer};
        % \node[block, below=of optimizer] (verifier) {Verifier};
        \node[draw, dashed, inner sep=0.3cm, rounded corners,
        % fit=(estimator)(feature)(optimizer)(verifier), label=above:{\textbf{Lucid}}] {};
        fit=(estimator)(feature)(optimizer)(sets)(tuner), label=below:{\textbf{Lucid}}] {};

        % Output block
        % \node[output-block, left=of verifier] (output) {Barrier certificate};
        \node[output-block, below=of optimizer] (output) {Barrier certificate};

@@

        % Arrows
        \draw[arrow] (input) -- (estimator);
        \draw[arrow] (input) -- (tuner);
        \draw[arrow] (input) -- (sets);
        \draw[arrow] (sets) -- (optimizer);
        \draw[arrow] (tuner) -- (estimator);
        \draw[arrow] (estimator) -- (feature);
        \draw[arrow] (feature) -- (optimizer);
        % \draw[arrow] (optimizer) -- (verifier);
        % \draw[arrow] (verifier) -- (output);
        \draw[arrow] (optimizer) -- (output);

@@

    \end{tikzpicture}
    \caption{General architecture of Lucid, highlighting its core components and their connections.}
    \label{fig:architecture}
\end{figure}

>>>>

\begin{figure}
    \centering
    \begin{tikzpicture}[node distance=0.6cm and 0.6cm,
            input-block/.style={rectangle, draw, fill=red!10, text width=2.5cm, minimum height=0.8cm, align=center, rounded corners},
            output-block/.style={rectangle, draw, fill=green!20, text width=2.6cm, minimum height=0.8cm, align=center, rounded corners},
            block/.style={rectangle, draw, fill=accent!20, text width=2.0cm, minimum height=0.8cm, align=center, rounded corners},
            arrow/.style={thick,->,>=latex},
            % tool/.style={rectangle, draw, fill=magenta!10, text width=3cm, minimum height=0.8cm, align=center, rounded corners},
            inv/.style={}
        ]

        % Input block
        \node[doc] (input) {Config\\(.yaml?)};

        % Lucid blocks
        \node[block, below=1cm of input] (estimator) {\Circled{2} Estimator};
        \node[block, right=of estimator] (tuner) {\Circled{3} Tuner};
        \node[block, left=of estimator] (sets) {\Circled{1} Safety Specification};
        \node[block, below=of estimator] (feature) {\Circled{4} Feature Map}; % / Barrier template
        \node[block, left=of feature] (optimizer) {\Circled{5} Solver};
        % \node[block, below=of optimizer] (verifier) {Verifier};
        \node[draw, dashed, inner sep=0.25cm, rounded corners,
        % fit=(estimator)(feature)(optimizer)(verifier), label=above:{\textbf{Lucid}}] {};
        fit=(estimator)(feature)(optimizer)(sets)(tuner), label=below:{\textbf{\lucid}}] {};

        % Output block
        % \node[output-block, left=of verifier] (output) {Barrier certificate};
        \node[output-block, below=of optimizer] (output) {Barrier Certificate};
        \node[output-block, below=.5cm of output] (prob) {Safety Probability};
        \node[] at ($(output)!0.5!(prob)$) (plus) {$+$};

@@

        % Arrows
        \draw[arrow] (input) -- (estimator);
        \draw[arrow] ($(input.south)+(0.0,-0.3)$) -| (tuner);
        \draw[arrow] ($(input.south)+(0.0,-0.3)$) -| (sets);
        \draw[arrow] (sets) -- (optimizer);
        \draw[arrow] (tuner) -- (estimator);
        \draw[arrow] (estimator) -- (feature);
        \draw[arrow] (feature) -- (optimizer);
        % \draw[arrow] (optimizer) -- (verifier);
        % \draw[arrow] (verifier) -- (output);
        \draw[arrow] (optimizer) -- (output);

@@

    \end{tikzpicture}
    \caption{General architecture of \lucid, highlighting its core components and their connections.}
    \label{fig:architecture}
\end{figure}

### 2025-07-26 docs/paper/figures/steps.tex
feat: add theory

        % Input block
        \node[input-block] (input) {Collect samples $\{(x_i, x^+_i)\}_{i=1}^N$};

        % Lucid blocks
        \node[block, right=of input] (estimator) {Fit the estimator $\hat{f}: \X \to \mathbb{R}^M$};
        \node[block, below=of estimator] (feature) {Construct the feature map $\phi: \X \to \mathbb{R}^M$};
        \node[block, below=of feature] (lattice) {Construct lattices on $\X, \X_0, \X_u$};
        \node[block, below=of lattice] (apply-lattice) {Apply the feature map to the lattices points};
        \node[block, below=of apply-lattice] (optimizer) {Optimize the LP};
        \node[draw, dashed, inner sep=0.3cm, rounded corners,
        fit=(estimator)(feature)(lattice)(apply-lattice)(optimizer), label=above:{\textbf{Lucid}}] {};

        % Output block
        \node[output-block, left=of optimizer] (output) {Barrier certificate};

>>>>

        % Input block
        \node[input-block] (input) {Collect samples $\{(x^i,a^i, x_+^i)\}_{i=1}^N$};

        % Lucid blocks
        \node[block, right=of input] (estimator) {Fit the estimator $\hat{f}\colon \X\times\A \to \X$};
        \node[block, below=of estimator] (feature) {Construct the feature map $\phi_M\colon \X\times\A \to \mathbb{R}^M$};
        \node[block, below=of feature] (lattice) {Construct lattices on $\X, \X_0, \X_u$};
        \node[block, below=of lattice] (apply-lattice) {Apply the feature map to the lattice points};
        \node[block, below=of apply-lattice] (optimizer) {Solve the LP};
        \node[draw, dashed, inner sep=0.3cm, rounded corners,
        fit=(estimator)(feature)(lattice)(apply-lattice)(optimizer), label=above:{\textbf{\lucid}}] {};

        % Output block
        \node[output-block, left=of optimizer] (output) {Barrier certificate\\+\\Safety probability};
